cardinalR: Generating interesting high-dimensional data structures

### 2025-06-24 paper-cardinalR.Rmd
Updated Pyramid function definitions

The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base. The pyramid is embedded in a $p$-dimensional space, with a tip at height zero and base at height $h$. The shape tapers linearly from the base dimensions $(l_x, l_y)$ to a smaller rectangular section with side lengths $(2r_t, 2r_t)$ at the tip. Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.

Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final dimension $x_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$:

$$
x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)
$$.

Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$:

$$
r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}, \quad
r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}
$$.

The first three coordinates are then defined as:

$$
x_1 \sim U(-r_x(z),\ r_x(z)), \quad
x_2 \sim U(-r_y(z),\ r_y(z)), \quad
x_3 \sim U(-r_x(z),\ r_x(z))
$$.

For $p > 3$, the remaining $p - 4$ dimensions (i.e., $x_4$ to $x_{p-1}$) are tapered toward zero with decreasing height and generated as:

$$
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
$$

This tapering ensures that the structure narrows in all directions as it approaches the tip.

@@

The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section. The structure is embedded in a $p$-dimensional space, with the tip located at height 0 and the base at height $h$. The triangle expands linearly in size from tip to base, with more points concentrated near the tip.

Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate \$x_p\$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:

$$
x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
$$

Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$:

$$
r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
$$

A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section:

$$
u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
$$

The first three coordinates (triangle plane) are then:

$$
x_1 = r(z)(1 - u - v), \quad
x_2 = r(z) \cdot u, \quad
x_3 = r(z) \cdot v
$$

For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip:

$$
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
$$

The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section. The structure is embedded in a $p$-dimensional space, with the tip located at height 0 and the base at height $h$. The triangle expands linearly in size from tip to base, with more points concentrated near the tip.

Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate $x_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$:

$$
x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
$$

Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$:

$$
r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
$$

A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section:

$$
u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
$$

The first three coordinates (triangle plane) are then:

$$
x_1 = r(z)(1 - u - v), \quad
x_2 = r(z) \cdot u, \quad
x_3 = r(z) \cdot v
$$

For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip:

$$
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
$$

The result is a tapered, high-dimensional triangle-based pyramid with more points near the apex and greater width at the base.

@@

The `gen_pyrstar(n, p, h, rb)` function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base. The pyramid extends vertically from height $z = 0$ (tip) to height $z = h$ (base), embedded in a $p$-dimensional space. The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.

Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The final coordinate $x_p$ encodes the height $z$:

$$
x_p = z \sim U(0, h).
$$

The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$:

$$
r(z) = r_b \cdot \left(1 - \frac{z}{h}\right).
$$

Each point is placed within a regular hexagon in the plane $(x_1, x_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor:

$$
\theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
r_{\text{point}} \sim \sqrt{U(0, 1)}.
$$

Then, the first two coordinates are:

$$
x_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta), \quad
x_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)
$$

For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip:

$$
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
$$

This results in a smooth, star-like pyramid in $p$-dimensional space, with a broad hexagonal base and narrow tip.

@@

The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p$-dimensional simplex using a chaotic attractor-like midpoint algorithm. The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.

Let $x_1, x_2, \dots, x_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution:

$$
T_0 \sim U(0, 1)^p.
$$

Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p$-dimensional simplex. At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$:

$$
T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}.
$$

This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex.

The final output is a $n \times p$ matrix where each row represents a point:

$$
X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}.
$$

>>>>

The `gen_pyrrect(n, p, h, l_vec, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a rectangular base. The pyramid is embedded in $p\text{-}D$, with a tip at height zero and base at height $h$. The shape tapers linearly from the base dimensions $(l_x, l_y)$ to a smaller rectangular section with side lengths $(2r_t, 2r_t)$ at the tip. Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.

Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points. The final dimension $X_p$ encodes the height of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$. Let $r_x(z)$ and $r_y(z)$ denote the half-widths of the rectangular cross-section at height $z$: $r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}$, $r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}$. The first three coordinates are then defined as: $X_1 \sim U(-r_x(z),\ r_x(z))$, $X_2 \sim U(-r_y(z),\ r_y(z))$, and $X_3 \sim U(-r_x(z),\ r_x(z))$. For $p > 3$, the remaining $p - 4$ dimensions (i.e., $X_4$ to $X_{p-1}$) are tapered toward zero with decreasing height and generated as: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1$. This tapering ensures that the structure narrows in all directions as it approaches the tip.

@@

The `gen_pyrtri(n, p, h, l, rt)` function generates a dataset of $n$ points forming a high-dimensional pyramid-like structure with a triangular cross-section. The structure is embedded in a $p\text{-}D$, with the tip located at height $0$ and the base at height $h$. The triangle expands linearly in size from tip to base, with more points concentrated near the tip.

Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points. The final coordinate $X_p$ encodes the height $z$ of each point and is drawn from an exponential distribution capped at $h$: $X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)$. Let $r(z)$ denote the scaling factor (distance from the origin to triangle vertices) at height $z$: $r(z) = r_t + (l - r_t) \cdot \frac{z}{h}$.

A point in the triangle at height $z$ is generated using barycentric coordinates $(u, v)$ to ensure uniform sampling within the triangular cross-section: $u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v$. The first three coordinates (triangle plane) are then: $X_1 = r(z)(1 - u - v)$, $X_2 = r(z) \cdot u$, and $X_3 = r(z) \cdot v$. For dimensions $j = 4, \dots, p-1$, values are tapered linearly toward zero near the tip: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right)$.

@@

The `gen_pyrstar(n, p, h, rb)` function generates a dataset of $n$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base. The pyramid extends vertically from height $z = 0$ (tip) to height $z = h$ (base), embedded in a $p$-dimensional space. The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.

Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points. The final coordinate $X_p$ encodes the height $z$: $X_p = z \sim U(0, h)$.

The radius at height $z$ scales linearly from zero (tip) to the base radius $r_b$: $r(z) = r_b \cdot \left(1 - \frac{z}{h}\right)$.

Each point is placed within a regular hexagon in the plane $(X_1, X_2)$, using a randomly chosen hexagon sector angle $\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}$ and a uniformly random radial scaling factor: $\theta \sim \text{Uniform sample from 6 hexagon angles}$,
$r_{\text{point}} \sim \sqrt{U(0, 1)}$. Then, the first two coordinates are: $X_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta)$, and $X_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)$. For dimensions $j = 3, \dots, p-1$, values are tapered toward zero near the tip: $X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right)$.

@@

The `gen_pyrholes(n, p)` function generates $n$ points embedded in a $p\text{-}D$ simplex using a chaotic attractor-like midpoint algorithm. The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a "Sierpinski-like pyramid" in high dimensions.

Let $X_1, X_2, \dots, X_p$ denote the coordinates of the generated points. The generation process begins with an initial point $T_0 \in [0, 1]^p$ drawn from a uniform distribution: $T_0 \sim U(0, 1)^p$.

Let $C_1, C_2, \dots, C_{p+1}$ denote the corner vertices of a $p\text{-}D$ simplex. At each iteration $i = 1, \dots, n$, a new point is computed by taking the midpoint between the previous point $T_{i-1}$ and a randomly selected vertex $C_k$: $T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}$. This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex. The final output is a $n \times p$ matrix where each row represents a point: $X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}$.

### 2025-06-24 paper-cardinalR.tex
Updated Pyramid function definitions

The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base. The pyramid is embedded in a \(p\)-dimensional space, with a tip at height zero and base at height \(h\). The shape tapers linearly from the base dimensions \((l_x, l_y)\) to a smaller rectangular section with side lengths \((2r_t, 2r_t)\) at the tip. Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.

Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final dimension \(x_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\):

\[
x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)
\].

Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\):

\[
r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}, \quad
r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}
\].

The first three coordinates are then defined as:

\[
x_1 \sim U(-r_x(z),\ r_x(z)), \quad
x_2 \sim U(-r_y(z),\ r_y(z)), \quad
x_3 \sim U(-r_x(z),\ r_x(z))
\].

For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(x_4\) to \(x_{p-1}\)) are tapered toward zero with decreasing height and generated as:

\[
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1.
\]

This tapering ensures that the structure narrows in all directions as it approaches the tip.

@@

The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section. The structure is embedded in a \(p\)-dimensional space, with the tip located at height 0 and the base at height \(h\). The triangle expands linearly in size from tip to base, with more points concentrated near the tip.

Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final coordinate \$x\_p\$ encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):

\[
x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
\]

Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\):

\[
r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
\]

A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section:

\[
u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
\]

The first three coordinates (triangle plane) are then:

\[
x_1 = r(z)(1 - u - v), \quad
x_2 = r(z) \cdot u, \quad
x_3 = r(z) \cdot v
\]

For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip:

\[
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
\]

The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section. The structure is embedded in a \(p\)-dimensional space, with the tip located at height 0 and the base at height \(h\). The triangle expands linearly in size from tip to base, with more points concentrated near the tip.

Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final coordinate \(x_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\):

\[
x_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right).
\]

Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\):

\[
r(z) = r_t + (l - r_t) \cdot \frac{z}{h}.
\]

A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section:

\[
u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v.
\]

The first three coordinates (triangle plane) are then:

\[
x_1 = r(z)(1 - u - v), \quad
x_2 = r(z) \cdot u, \quad
x_3 = r(z) \cdot v
\]

For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip:

\[
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right).
\]

The result is a tapered, high-dimensional triangle-based pyramid with more points near the apex and greater width at the base.

@@

The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \$n\$ points forming a high-dimensional pyramid-shaped structure with a hexagonal base. The pyramid extends vertically from height \(z = 0\) (tip) to height \(z = h\) (base), embedded in a \(p\)-dimensional space. The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.

Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The final coordinate \(x_p\) encodes the height \(z\):

\[
x_p = z \sim U(0, h).
\]

The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\):

\[
r(z) = r_b \cdot \left(1 - \frac{z}{h}\right).
\]

Each point is placed within a regular hexagon in the plane \((x_1, x_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor:

\[
\theta \sim \text{Uniform sample from 6 hexagon angles}, \quad
r_{\text{point}} \sim \sqrt{U(0, 1)}.
\]

Then, the first two coordinates are:

\[
x_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta), \quad
x_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)
\]

For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip:

\[
x_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right).
\]

This results in a smooth, star-like pyramid in \(p\)-dimensional space, with a broad hexagonal base and narrow tip.

@@

The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\)-dimensional simplex using a chaotic attractor-like midpoint algorithm. The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.

Let \(x_1, x_2, \dots, x_p\) denote the coordinates of the generated points. The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution:

\[
T_0 \sim U(0, 1)^p.
\]

Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\)-dimensional simplex. At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\):

\[
T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}.
\]

This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex.

The final output is a \(n \times p\) matrix where each row represents a point:

\[
X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}.
\]

>>>>

The \texttt{gen\_pyrrect(n,\ p,\ h,\ l\_vec,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a rectangular base. The pyramid is embedded in \(p\text{-}D\), with a tip at height zero and base at height \(h\). The shape tapers linearly from the base dimensions \((l_x, l_y)\) to a smaller rectangular section with side lengths \((2r_t, 2r_t)\) at the tip. Points are distributed more densely near the tip, simulating a natural skew toward smaller height values.

Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points. The final dimension \(X_p\) encodes the height of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\). Let \(r_x(z)\) and \(r_y(z)\) denote the half-widths of the rectangular cross-section at height \(z\): \(r_x(z) = r_t + (l_x - r_t) \cdot \frac{z}{h}\), \(r_y(z) = r_t + (l_y - r_t) \cdot \frac{z}{h}\). The first three coordinates are then defined as: \(X_1 \sim U(-r_x(z),\ r_x(z))\), \(X_2 \sim U(-r_y(z),\ r_y(z))\), and \(X_3 \sim U(-r_x(z),\ r_x(z))\). For \(p > 3\), the remaining \(p - 4\) dimensions (i.e., \(X_4\) to \(X_{p-1}\)) are tapered toward zero with decreasing height and generated as: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right), \quad \text{for } j = 4, \dots, p-1\). This tapering ensures that the structure narrows in all directions as it approaches the tip.

@@

The \texttt{gen\_pyrtri(n,\ p,\ h,\ l,\ rt)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-like structure with a triangular cross-section. The structure is embedded in a \(p\text{-}D\), with the tip located at height \(0\) and the base at height \(h\). The triangle expands linearly in size from tip to base, with more points concentrated near the tip.

Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points. The final coordinate \(X_p\) encodes the height \(z\) of each point and is drawn from an exponential distribution capped at \(h\): \(X_p = z \sim \min\left(\text{Exp}(\lambda = 2/h),\ h\right)\). Let \(r(z)\) denote the scaling factor (distance from the origin to triangle vertices) at height \(z\): \(r(z) = r_t + (l - r_t) \cdot \frac{z}{h}\).

A point in the triangle at height \(z\) is generated using barycentric coordinates \((u, v)\) to ensure uniform sampling within the triangular cross-section: \(u, v \sim U(0, 1), \quad \text{if } u + v > 1: u \leftarrow 1 - u,\ v \leftarrow 1 - v\). The first three coordinates (triangle plane) are then: \(X_1 = r(z)(1 - u - v)\), \(X_2 = r(z) \cdot u\), and \(X_3 = r(z) \cdot v\). For dimensions \(j = 4, \dots, p-1\), values are tapered linearly toward zero near the tip: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h}\right)\).

@@

The \texttt{gen\_pyrstar(n,\ p,\ h,\ rb)} function generates a dataset of \(n\) points forming a high-dimensional pyramid-shaped structure with a hexagonal base. The pyramid extends vertically from height \(z = 0\) (tip) to height \(z = h\) (base), embedded in a \(p\)-dimensional space. The distribution concentrates more points near the base, and the hexagonal spread tapers toward the tip.

Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points. The final coordinate \(X_p\) encodes the height \(z\): \(X_p = z \sim U(0, h)\).

The radius at height \(z\) scales linearly from zero (tip) to the base radius \(r_b\): \(r(z) = r_b \cdot \left(1 - \frac{z}{h}\right)\).

Each point is placed within a regular hexagon in the plane \((X_1, X_2)\), using a randomly chosen hexagon sector angle \(\theta \in \{0, \frac{\pi}{3}, \frac{2\pi}{3}, \pi, \frac{4\pi}{3}, \frac{5\pi}{3}\}\) and a uniformly random radial scaling factor: \(\theta \sim \text{Uniform sample from 6 hexagon angles}\),
\(r_{\text{point}} \sim \sqrt{U(0, 1)}\). Then, the first two coordinates are: \(X_1 = r(z) \cdot r_{\text{point}} \cdot \cos(\theta)\), and \(X_2 = r(z) \cdot r_{\text{point}} \cdot \sin(\theta)\). For dimensions \(j = 3, \dots, p-1\), values are tapered toward zero near the tip: \(X_j \sim U(-0.1, 0.1) \cdot \left(1 - \frac{z}{h} \right)\).

@@

The \texttt{gen\_pyrholes(n,\ p)} function generates \(n\) points embedded in a \(p\text{-}D\) simplex using a chaotic attractor-like midpoint algorithm. The result is a fractal-like structure that reveals holes or gaps in the data cloud, forming a ``Sierpinski-like pyramid'' in high dimensions.

Let \(X_1, X_2, \dots, X_p\) denote the coordinates of the generated points. The generation process begins with an initial point \(T_0 \in [0, 1]^p\) drawn from a uniform distribution: \(T_0 \sim U(0, 1)^p\).

Let \(C_1, C_2, \dots, C_{p+1}\) denote the corner vertices of a \(p\text{-}D\) simplex. At each iteration \(i = 1, \dots, n\), a new point is computed by taking the midpoint between the previous point \(T_{i-1}\) and a randomly selected vertex \(C_k\): \(T_i = \frac{1}{2}(T_{i-1} + C_k), \quad C_k \in \{C_1, \dots, C_{p+1}\}\). This recursive midpoint rule generates self-similar patterns with systematic voids (holes) between clusters of points. The points remain bounded inside the convex hull of the simplex. The final output is a \(n \times p\) matrix where each row represents a point: \(X = \{T_1, T_2, \dots, T_n\}, \quad X \in \mathbb{R}^{n \times p}\).
